\chapter{Methods (5-7 pages)}


\section{Design}
%-------------------------------------------%

Design details of your project. This section can include various descriptions, figures, diagrams, wireframes, etc that are driving your project implementation.

%-------------------------------------------%



\section{Frameworks}
%-------------------------------------------%

Statement and descriptions of systems, languages, frameworks, APIs, and tools used in the development of your project.

%-------------------------------------------%



\section{Algorithms}
%-------------------------------------------%

Descriptions and pseudocode of important algorithms in use, relevant, or to be implemented as part of your project.

%-------------------------------------------%



\section{Analytical Methods}
%-------------------------------------------%

Descriptions and processes of any analytical methods used as part of your capstone project. This section is most applicable for projects with a data analysis focus or component. Quantitative and qualitative methods.

%-------------------------------------------%



\section{Features}
%-------------------------------------------%

A brief overview of primary features you plan to implement.

%-------------------------------------------%



\section{Test Plan}
%-------------------------------------------%

A brief overview of what and how you plan to test your project codebase.

%-------------------------------------------%



\section{Criteria and Constraints}
%-------------------------------------------%

Project development must meet desired needs within realistic constraints. Many times, constraints are interrelated. Cover the appropriate criteria and constraints related to your capstone. Some example criteria and constraints are health and safety, environmental, political, social, manufacturability, sustainability, economical, and ethical.

%-------------------------------------------%